Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-07-17 draft.tex
Add par. and comments to detector effects

The determination of detector acceptance is generally a difficult task. Here, we will show a systematic
method to estimate its effects based on both the method of moments, and MC simulations of the detector.\\

As previous, we use a PDF $P$ decomposed into $N$ angular observables $\lbrace S_k\rbrace$,
\begin{equation}
    P \equiv P(\vec\nu,\vec\theta | \lbrace S_k \rbrace)\,.
\end{equation}
We denote the estimator of an
observable $S_k$ as $\est{S_k}$, which is affected by the detector acceptance. We assume now that there is a linear relation
between the true angular observables $\lbrace S_k^\text{true}\rbrace$, and the estimators $\lbrace \est{S_k}\rbrace$,
\begin{equation}
    \label{eq:linear-assumption}
    \est{S_i} = \sum_{i = 1}^N M_{i j} S_j^\text{true}\,.
\end{equation}
In that case, the matrix $M_{ij}$ can be fully determined through the following steps.
\begin{enumerate}
    \item For a fixed $j$, let $S_j^\text{true} = 1$, and $S_k^\text{true} = 0\,\forall k\neq j$.
    \item Draw simulated events $\lbrace x_t^{(k)}\rbrace = \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ (dubbed ``true'') from the PDF $P_j$,
        \begin{equation}
            P_j \equiv P(\vec\nu,\vec\theta | \lbrace S_j = 1, S_{k\neq j} = 0\rbrace)\,.
        \end{equation}
    \item Use the true events as input for the MC simulation of the detector, in order to obtain the detector
        events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,\vec\theta^{(k)}_d)\rbrace$.
    \item Use the method of moments to determine the uncorrected observables $\lbrace \est{S_k} \rbrace$, based on the detector
        events $\lbrace x^{(k)}_d\rbrace$.
    \item By virtue of \refeq{linear-assumption} and step 1, the $j$-th column of $M$ has been determined via
        \begin{equation}
            M_{ij} = \est{S_i}\,.
        \end{equation}
    \item Repeat the above steps for all values of $j$ to determine all columns of $M$.
\end{enumerate}

>>>>

The determination of detector acceptance is generally a difficult
task. Here, we will show a systematic method to estimate its effects
based on both the method of moments, and MC simulations of the
detector.

\fred{In the ideal case, one would have an explicit probabilistic
  model of the detector acceptance effects and could thus write down
  the full forward model of the measured events arise. In practice,
  this would require much more CPU time than available, and one is
  therefore forced to simplify the model. The standard approach is to
  generate \emph{true} particle events $\lbrace x_t^{(k)}\rbrace =
  \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ from a PDF
  assumed to describe the bare physical process, and to propagate
  those particles through a detailed simulation of the detector. The
  observable traces that the particles leave in the detector are fed
  into reconstruction algorithms resulting in the \emph{detector}
  events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,
  \vec\theta^{(k)}_d)\rbrace$.

[Where does the reconstruction uncertainty enter? Below we just
  account for systematic shifts but our estimate of the uncertainty of
  one $S_k$ would be only the statistical uncertainty from the finite
  sample size. How about the extra correlation between different $S_k$
  that is induced?] }\\

As before, we assume a PDF $P$ can be decomposed into $N$ parts with angular observables $\lbrace S_k\rbrace$,
\begin{equation}
    P \equiv P(\vec\nu,\vec\theta | \lbrace S_k \rbrace)\,.
\end{equation}
We denote the estimator of an
observable $S_k$ as $\est{S_k}$, which is affected by the detector acceptance. We assume now that there is a linear relation
between the true angular observables $\lbrace S_k^\text{true}\rbrace$, and the estimators $\lbrace \est{S_k}\rbrace$,
\fred{Do we need $S_k^\text{true}$? Isn't it just $S_k$?}
\begin{equation}
    \label{eq:linear-assumption}
    \est{S_i} = \sum_{i = 1}^N M_{i j} S_j^\text{true}\,.
\end{equation}
In that case, the matrix $M_{ij}$ can be fully determined through the following steps.
\begin{enumerate}
    \item For a fixed $j$, let $S_j^\text{true} = 1$, and $S_k^\text{true} = 0\,\forall k\neq j$.
    \item Draw simulated events $\lbrace x_t^{(k)}\rbrace = \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ (dubbed ``true'') from the PDF $P_j$,
        \begin{equation}
            P_j \equiv P(\vec\nu,\vec\theta | \lbrace S_j = 1, S_{k\neq j} = 0\rbrace)\,.
        \end{equation}
    \item Use the true events as input for the MC simulation of the detector, in order to obtain the detector
        events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,\vec\theta^{(k)}_d)\rbrace$.
    \item Use the method of moments to determine the uncorrected observables $\lbrace \est{S_k} \rbrace$, based on the detector
        events $\lbrace x^{(k)}_d\rbrace$.
    \item By virtue of \refeq{linear-assumption} and step 1, the $j$-th column of $M$ has been determined via
        \begin{equation}
            M_{ij} = \est{S_i}\,.
        \end{equation}
        \fred{I wonder if a small sample size also leads to artificial
          nonzero off-diagonal elements of $M$}
    \item Repeat the above steps for all values of $j$ to determine all columns of $M$.
\end{enumerate}
